Dancing on Water: The Choreography of Sulfur Dioxide Adsorption to Aqueous Surfaces

### 2011-07-05 Conclusions.tex
Final draft

The simulations described herein show that gaseous \suldiox~quickly adsorbs to the water surface and continues to bind until a complete surface coverage is reached. Surface waters reorient in the presence of adsorbed \suldiox. The waters at the interface of a neat-water surface tend to lay flatter to the surface than when a saturating layer of \suldiox~is present. The waters interacting with the layer of adsorbed \suldiox~orient more perpendicularly to the interface, and further expose their ``free-OH'' uncoupled bonds for interactions with \suldiox, and hydrate complex formation. Furthermore, we have found that surface waters underneath a blanket layer of adsorbed \suldiox~will penetrate further into the gas phase, allowing for greater mobility of waters away from the aqueous bulk in the presence of \suldiox.

@@

These studies provide a starting point for future studies in this area that seek to understand how gases of different concentrations and chemical composition adsorb and transit across an aqueous/air interface. Obtaining such knowledge will be invaluable for understanding many environmental aerosol and land water systems where gaseous uptake at a water surface does not conform to expectations. In such cases, surface complexes are often invoked with little direct evidence of the formation of such complexes. Investigations of these and related low-temperature effects will be forthcoming in future publications.

>>>>

Our simulations show that gaseous \suldiox~quickly adsorbs to the water surface and continues to bind until a complete surface coverage is reached. Surface waters reorient in the presence of adsorbed \suldiox. The waters at the interface of a neat-water surface tend to lay flatter to the surface than when a saturating layer of \suldiox~is present. The waters interacting with the layer of adsorbed \suldiox~orient more perpendicularly to the interface, and further expose their ``free-OH'' uncoupled bonds for interactions with \suldiox, and hydrate complex formation. Furthermore, we have found that surface waters underneath a blanket layer of adsorbed \suldiox~will penetrate further into the gas phase, allowing for greater mobility of waters away from the aqueous bulk in the presence of \suldiox.

@@

These studies provide a starting point for future studies in this area that seek to understand how gases of different concentrations and chemical composition adsorb and transit across an aqueous/air interface. Obtaining such knowledge will be invaluable for understanding many environmental aerosol and land water systems where gaseous uptake at a water surface does not conform to expectations.\cite{Jayne1990,Yang2002,Worsnop1989,Boniface2000} Investigations of these and related low-temperature effects will be forthcoming in future publications.

% In such cases, surface complexes are often invoked with little direct evidence of the formation of such complexes.

### 2011-07-05 DensityResults.tex
Final draft

The saturated solution simulation results indicate that under the high concentration of \suldiox, the \suldiox~accumulation at the surface is increased. However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox. The added concentration of \suldiox~creates a layer of molecules bound to the top of the water surface. The center of the \suldiox~density distribution is further into the gas phase than for the neat-water surface with a single-\suldiox~molecule. Because of the limitations of the classical model in accurately reproducing the first hydration shell around the \suldiox,\cite{Baer2010} we do not draw specific conclusions about the hydrated surface complex's geometry. However, the surface affinity of the hydrated \suldiox~is well reproduced by both the classical and ab initio methods.

>>>>

The saturated solution simulation results indicate that under the high concentration of \suldiox, the \suldiox~accumulation at the surface is increased. However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox. The added concentration of \suldiox~creates a layer of molecules bound to the top of the water surface. The center of the \suldiox~density distribution is further into the gas phase than for the neat-water surface with a single-\suldiox~molecule. Because of the limitations of the classical model in accurately reproducing the first hydration shell around the \suldiox, we do not draw specific conclusions about the hydrated surface complex's geometry. However, the surface affinity of the hydrated \suldiox~is well reproduced by both the classical and ab initio methods.\cite{Baer2010}

### 2011-07-05 EquilibratedOrientation.tex
Final draft

%The histograms show overall similarities for both systems in their shapes and intensities from approx. 0\angs~and below. The presence of a saturated layer of adsorbed \suldiox~molecules, however, alters the water orientation, but not necessarily the orientational depth of the interface. In both $\theta$ distributions, the orientations of waters in the bulk region are isotropic until 5\angs~below the surface location. At 5\angs~below the surface the water molecules begin to orient with their bisectors within the plane of the interface (perpendicular to the surface normal, $\cos(\theta)\approx 0$). The corresponding location in the plot of $\phi$ shows that those molecules are also mostly flat to the surface ($\cos(\phi)\approx 1$). Moving further out from the bulk and into the gas phase, the distributions show $\cos(\theta)$ increasing. Waters further out from the bulk have fewer bonding interactions, and orient with their hydrogens more towards the gas phase. The bisector pointing further into the gas phase leads to isotropy in the values of $\phi$.

The noise in the neat-water plots of Figure \ref{fig:water-orientation} above 5\angs~(manifested as disconnected points of intensity throughout the range of orientations) is a result of fewer waters venturing beyond those extents and thus less data far from the water surface location. This is one indication that waters near a layer of adsorbed \suldiox~venture further above the interface relative to the low \suldiox~concentration, where they can have more \suldiox~interaction with the adsorbed gas molecules. These results show that the interactions with neighboring \suldiox~molecules allow the waters above the surface to orient more perpendicularly to the interface. This is consistent with our recent experimental VSFS studies which showed evidence for the reorienting behavior of water due to the \suldiox~interactions with the topmost surface waters.\cite{Ota2011}

The distribution of $\cos(\phi)$ is more sharply defined (i.e. less isotropic) for the neat-water system than for the saturated one. Waters on the neat surface lie flatter, whereas the presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface. The $\phi$ distributions quickly become isotropic above the surface as the bisectors orient more perpendicularly, and below the surface as the bulk water loses any orientational preference.

>>>>

%The histograms show overall similarities for both systems in their shapes and intensities from approximately 0\angs~and below. The presence of a saturated layer of adsorbed \suldiox~molecules, however, alters the water orientation, but not necessarily the orientational depth of the interface. In both $\theta$ distributions, the orientations of waters in the bulk region are isotropic until 5\angs~below the surface location. At 5\angs~below the surface the water molecules begin to orient with their bisectors within the plane of the interface (perpendicular to the surface normal, $\cos(\theta)\approx 0$). The corresponding location in the plot of $\phi$ shows that those molecules are also mostly flat to the surface ($\cos(\phi)\approx 1$). Moving further out from the bulk and into the gas phase, the distributions show $\cos(\theta)$ increasing. Waters further out from the bulk have fewer bonding interactions, and orient with their hydrogens more towards the gas phase. The bisector pointing further into the gas phase leads to isotropy in the values of $\phi$.

The noise in the neat-water plots of Figure \ref{fig:water-orientation} above 5\angs~(manifested as disconnected points of intensity throughout the range of orientations) is a result of fewer waters venturing beyond those extents and thus less data far from the water surface location. This is one indication that waters near a layer of adsorbed \suldiox~venture further above the water surface location relative to the low \suldiox~concentration, where they can have more \suldiox~interaction with the adsorbed gas molecules. These results show that the interactions with neighboring \suldiox~molecules allow the waters above the water surface location to orient more perpendicularly to the interface. This is consistent with our recent experimental VSFS studies which showed evidence for the reorienting behavior of water due to the \suldiox~interactions with the topmost surface waters.\cite{Ota2011}

The distribution of $\cos(\phi)$ is more sharply defined (i.e. less isotropic) for the neat-water system than for the saturated one. Waters on the neat surface lie flatter, whereas the presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface. The $\phi$ distributions quickly become isotropic above the surface location as the bisectors orient more perpendicularly, and below the surface as the bulk water loses any orientational preference.

### 2011-07-05 TransitOrientation.tex
Lots of small corrections (typos!!) to draft 8 handed back for polishing.
Final draft

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/angle-cartoons/so2-surface.png}
		\caption{(A) \suldiox~molecules above and below both neat-water and saturated surfaces orient with the bisectors perpendicular to the interface. The \suldiox~sulfur points down towards the aqueous phase. (B) In the neat-water system, a transiting \suldiox~molecule located below the surface briefly reorients with one S-O bond more towards the interface above, and the other bond pointing further down into the water bulk.}
		\label{fig:so2-surface-angles}
	\end{center}
\end{figure}

@@

 With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ for \suldiox~would be isotropic relative to the reference axis. This is the case in both systems, with only a few exceptions. In both systems the $\cos(\phi)$ profiles exhibit mostly isotropic distributions above 0\angs, with several regions of lighter coloration interspersed, but without a clearly formed orientational trend. At the neat-water surface and just below, from 0\angs~to -5\angs~(in the water phase), the $\theta$ profile broadens down to $\cos(\theta)=0.5$, coinciding with a clear trend in the profile of $\phi$ near $\cos(\phi)=0$ shown in dark red at the bottom of the plot. This corresponds to \suldiox~inclined up to 60 degrees from the surface normal. The result is that \suldiox~in this region shows a preference for one of the S-O bonds to point more towards the aqueous bulk and the other to point away towards the interface. Figure \ref{fig:so2-surface-angles}b depicts the orientation of the \suldiox~slightly below the water surface. This is in contrast to \suldiox~molecules above the surface, oriented more perpendicularly. The behavior is likely due to the interaction between the waters and one of the S-O bonds leading to a higher solvation than above the water surface. As the \suldiox~is solvated by more highly-coordinated bulk water, the S-O bonds experience less equal interaction environments. Baer et al. noted that their force field model for the \suldiox~does not reproduce well the first hydration shell geometries,\cite{Baer2010} so conclusions regarding the specific interactions and hydrate geometries between the \suldiox~and \wat~cannot be made here. It is notable that the same reorientation does not occur as strongly in the saturated system. The presence of the adsorbed \suldiox~layer apparently decreases the reorienting behavior likely because of the disrupting effect the higher \suldiox~concentration has on the water interactions at the surface. %Below this 5\angs~region under the surface, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.

>>>>

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/angle-cartoons/so2-surface.png}
		\caption{(A) \suldiox~molecules above and below both neat-water and saturated water surface location orient with the bisectors perpendicular to the interface. The \suldiox~sulfur points down towards the aqueous phase. (B) In the neat-water system, a transiting \suldiox~molecule located below the water surface briefly reorients with one S-O bond more towards the interface above, and the other bond pointing further down into the water bulk.}
		\label{fig:so2-surface-angles}
	\end{center}
\end{figure}

@@

 With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ for \suldiox~would be isotropic relative to the reference axis. This is the case in both systems, with only a few exceptions. In both systems the $\cos(\phi)$ profiles exhibit mostly isotropic distributions above 0\angs, with several regions of lighter coloration interspersed, but without a clearly formed orientational trend. At the neat-water surface and just below, from 0\angs~to -5\angs~(in the water phase), the $\theta$ profile broadens down to $\cos(\theta)=0.5$, coinciding with a clear trend in the profile of $\phi$ near $\cos(\phi)=0$ shown in dark red at the bottom of the plot. This corresponds to \suldiox~inclined up to 60 degrees from the surface normal. The result is that \suldiox~in this region shows a preference for one of the S-O bonds to point more towards the aqueous bulk and the other to point away towards the interface. Figure \ref{fig:so2-surface-angles}B depicts the orientation of the \suldiox~slightly below the water surface location. This is in contrast to \suldiox~molecules above the water surface location, oriented more perpendicularly. The behavior is likely due to the interaction between the waters and one of the S-O bonds leading to a higher solvation than above the water surface location. As the \suldiox~is solvated by more highly-coordinated bulk water, the S-O bonds experience less equal interaction environments. Baer et al. noted that their force field model for the \suldiox~does not reproduce well the first hydration shell geometries,\cite{Baer2010} so conclusions regarding the specific interactions and hydrate geometries between the \suldiox~and \wat~cannot be made here. It is notable that the same reorientation does not occur as strongly in the saturated system. The presence of the adsorbed \suldiox~layer apparently decreases the reorienting behavior likely because of the disrupting effect the higher \suldiox~concentration has on the water interactions in the interfacial region. %Below this 5\angs~region under the surface, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.
